
% ████████████████████████████████████████████████████████████████████████████
% █░░░░░░░░░░░░░░███░░░░░░░░░░█░░░░░░░░░░░░░░███░░░░░░█████████░░░░░░░░░░░░░░█
% █░░▄▀▄▀▄▀▄▀▄▀░░███░░▄▀▄▀▄▀░░█░░▄▀▄▀▄▀▄▀▄▀░░███░░▄▀░░█████████░░▄▀▄▀▄▀▄▀▄▀░░█
% █░░▄▀░░░░░░▄▀░░███░░░░▄▀░░░░█░░▄▀░░░░░░▄▀░░███░░▄▀░░█████████░░▄▀░░░░░░▄▀░░█
% █░░▄▀░░██░░▄▀░░█████░░▄▀░░███░░▄▀░░██░░▄▀░░███░░▄▀░░█████████░░▄▀░░██░░▄▀░░█
% █░░▄▀░░░░░░▄▀░░░░███░░▄▀░░███░░▄▀░░░░░░▄▀░░░░█░░▄▀░░█████████░░▄▀░░██░░▄▀░░█
% █░░▄▀▄▀▄▀▄▀▄▀▄▀░░███░░▄▀░░███░░▄▀▄▀▄▀▄▀▄▀▄▀░░█░░▄▀░░█████████░░▄▀░░██░░▄▀░░█
% █░░▄▀░░░░░░░░▄▀░░███░░▄▀░░███░░▄▀░░░░░░░░▄▀░░█░░▄▀░░█████████░░▄▀░░██░░▄▀░░█
% █░░▄▀░░████░░▄▀░░███░░▄▀░░███░░▄▀░░████░░▄▀░░█░░▄▀░░█████████░░▄▀░░██░░▄▀░░█
% █░░▄▀░░░░░░░░▄▀░░█░░░░▄▀░░░░█░░▄▀░░░░░░░░▄▀░░█░░▄▀░░░░░░░░░░█░░▄▀░░░░░░▄▀░░█
% █░░▄▀▄▀▄▀▄▀▄▀▄▀░░█░░▄▀▄▀▄▀░░█░░▄▀▄▀▄▀▄▀▄▀▄▀░░█░░▄▀▄▀▄▀▄▀▄▀░░█░░▄▀▄▀▄▀▄▀▄▀░░█
% █░░░░░░░░░░░░░░░░█░░░░░░░░░░█░░░░░░░░░░░░░░░░█░░░░░░░░░░░░░░█░░░░░░░░░░░░░░█
% ████████████████████████████████████████████████████████████████████████████

\begin{thebibliography}{}

    \bibitem{mrison}  Mark Rison Z80 page for !CPC. \\
    {\tt \small http://www.acorn.co.uk/$\sim$mrison/en/cpc/tech.html}

    \bibitem{yaze} YAZE (Yet Another Z80 Emulator). This is a CPM emulator by Frank Cringle. It emulates almost every undocumented flag, very good emulator. Also includes a very good instruction exerciser and is released under the GPL. \\
    {\tt \small ftp://ftp.ping.de/pub/misc/emulators/yaze-1.10.tar.gz} \\
    Note: the instruction exerciser zexdoc/zexall does not test I/O instructions and not all normal instructions (for instance LD A,(IX+n) is tested, but not with different values of n, just n=1, values above 128 (LD A,(IX-n) are not tested) but it still gives a pretty good idea of how well a simulated Z80 works.

    \bibitem{thomas} Z80 Family Official Support Page by Thomas Scherrer. Very good -- your one-stop Z80 page. \\
    {\tt \small http://www.geocities.com/SiliconValley/Peaks/3938/z80\_home.htm}

    \bibitem{speccy_faq} Spectrum FAQ technical information. \\
    {\tt \small http://www.worldofspectrum.org/faq/}

    \bibitem{gerton} Gerton Lunter's Spectrum emulator (Z80). In the package there is a file TECHINFO.DOC, which contains a lot of interesting information. Note that the current version can only be unpacked in Windows. \\
    {\tt \small ftp://ftp.void.jump.org/pub/sinclair/emulators/pc/dos/z80-400.zip}

    \bibitem{mostek} Mostek Z80 Programming Manual -- a very good reference to the Z80.

    \bibitem{datasheet} Z80 Product Specification, from MSX2 Hardware Information. \\
    {\tt \small http://www.hardwareinfo.msx2.com/pdf/Zilog/z80.pdf}

    \bibitem{zx_next} ZX Spectrum Next information. \\
    {\tt \small \url{https://wiki.specnext.dev/}}

\end{thebibliography}

\pagebreak
\IntentionallyEmpty
\pagebreak