\section{Palette}
\label{zx_next_palette}

Next greatly enhances ZX Spectrum video capabilities by offering several new ways to draw graphics on a screen. We'll see how to program each in later chapters, but let's check common behaviour first - colour management.

To draw a pixel on a screen, we need to set its colour as data in memory. Next shares implementation to other contemporary 8-bit computers - all possible colours are stored together in a palette, as an array of RGB values, and each pixel is simply an index into this array. This approach requires less memory and allows creating efficient effects such as fade to/from black, transitions from day to night, water animations etc.


\subsection{Palette Selection}

Each graphics mode or layer has not one but two palettes, each of which can be changed independently. Of course, only one of two can be active at any given time for each mode. The other can be initialized with alternate colours and can be quickly activated to achieve colour animation effects.

Active palette is set with \PortLink{Enhanced ULA Control Register}{43} for ULA, Layer 2 and Sprites and \PortLink{Tilemap Control Register}{6B} for Tilemap.


\subsection{Palette Editing}

Next palettes have 256 colours. All are initialized with default values, so they are usable out of the box. But it's also possible to change every colour. Regardless of the palette, the procedure to read or write colours is:

\begin{enumerate}[topsep=1pt,itemsep=1pt]
	\item \PortLink{Enhanced ULA Control Register}{43} selects palette which colours you want to edit
	\item \PortLink{Palette Index Register}{40} selects colour index that will be read or written
	\item \PortLink{Palette Value Register}{41} or \PortLink{Enhanced ULA Palette Extension}{44} reads or writes data for selected colour
\end{enumerate}

When writing colours, we can chose to automatically increment colour indexes after each write. Bit 7 of \PortLink{Enhanced ULA Control Register}{43} is used for that purpose. This works the same for both write registers (\MemAddr{41} and \MemAddr{44}). Colour RGB values can either be 8-bit {\tt RRRGGGBB}, or 9-bit {\tt RRRGGGBBB} values. Use \PortLink{Palette Value Register}{41} for 8-bit and \PortLink{Enhanced ULA Palette Extension}{44} for 9-bit.

Note: \PortLink{Enhanced ULA Control Register}{43} has two roles when working with palettes - it selects the active palette for display (out of two available - only for ULA, Layer 2 and Sprites) and selects palette for editing (for all layers, including Tilemap). Therefore care needs to be taken when updating colour entries to avoid accidentally changing the active palette for display at the same time. Depending on our program, we may first need to read the value and then only change bits affecting the palette for editing to ensure the rest of the data remains unaffected.


\subsection{8 Bit Colours}

8-bit colours are stored as {\tt RRRGGGBB} values with 3 bits per red and green and 2 bits per blue component. Each colour is therefore stored as a single byte. \PortLink{Palette Value Register}{41} is used to read or write the value.

Here's a reusable subroutine for copying {\tt B} number of colours stored as a contiguous block in memory addressed by {\tt HL} register, starting at the currently selected colour index:

\begin{tcblisting}{}
Copy8BitPalette:
	LD A, (HL)                ; Load RRRGGGBB into A
	INC HL                    ; Increment to next colour entry
	NEXTREG &41, A            ; Send colour data to Next HW
	DJNZ Copy8BitPalette      ; Repeat until B=0
\end{tcblisting}

And we'd call the subroutine with something like:

\begin{tcblisting}{}
	NEXTREG &43, %00010000    ; Auto increment, Layer 2 first palette for read/write
	NEXTREG &40, 0            ; Start copying into index 0
	LD HL, palette            ; Address to copy RRRGGGBB values from
	LD B, 255                 ; Copy 255 colours
	CALL Copy8BitPalette
\end{tcblisting}

Note: we could also use DMA to transfer all the bytes. I won't show it here, but feel free to implement it as an excercise - see section \ref{zx_next_dma} for details on programming the DMA.


\subsection{9 Bit Colours}

With 9 bits per colour, each RGB component uses full 3 bits, thus greatly increasing the available colour gamut. However, each colour needs 2 bytes in memory instead of 1. To read or write we use \PortLink{Enhanced ULA Palette Extension}{44} register instead of \MemAddr{41}. It works similarly to \MemAddr{41} except that each colour requires two writes: first one stores {\tt RRRGGGBB} part and second least significant bit of blue component. Subroutine for copying 9-bit colours:

\begin{tcblisting}{}
Copy9BitPalette:
	LD A, (HL)                ; Load RRRGGGBB into A
	INC HL                    ; Increment to next byte
	NEXTREG &44, A            ; Send colour data to Next HW
	LD A, (HL)                ; Load remaining byte with LSB of B component into A
	INC HL                    ; Increment to next colour entry
	NEXTREG &44, A            ; Send colour data to Next HW and increment index
	DJNZ Copy9BitPalette      ; Repeat until B=0
\end{tcblisting}

Note: subroutine requires that colours are stored in 2 bytes with first containing {\tt RRRGGGBB} part and second least significant bit of blue. Which is how typically drawing programs store a 9-bit palette anyways. The code for calling this subroutine is exactly the same as for the 8-bit colours above.


\subsection{Palette Registers}
\label{zx_next_palette_registers}

\subsubsection{\PortTitle{Palette Index Register}{40}}

\begin{NextPort}
	\PortBits{7-0}
		\PortDesc{Reads or writes palette colour index to be manipulated}
\end{NextPort}

Writing an index {\tt 0-255} associates it with colour set through \PortLink{Palette Value Register}{41} or \PortLink{Enhanced ULA Palette Extension}{44} of currently selected pallette in \PortLink{Enhanced ULA Control Register}{43}. Write also resets value of \PortLink{Enhanced ULA Palette Extension}{44} so next write will occur for first colour of the palette.

While Tilemap, Layer 2 and Sprites palettes use all 256 distinct colours (with some caveats, as described in specific chapters), ULA modes work like this:

\begin{quote}
	\small

	\NewDocumentEnvironment{PalIndexTable}{ +b }{
		\begin{tabularx}{\linewidth}{@{}p{1.5cm}X}
			Index & Colours \\
			#1
		\end{tabularx}
		\vspace*{-1ex}
	}{}

	\NewDocumentCommand{\PalIndexEntry}{ m m }{{\tt #1} & #2 \\}

	\textbf{Classic ULA}

	\begin{PalIndexTable}
		\PalIndexEntry{~0-7}{Ink}
		\PalIndexEntry{~8-15}{Bright ink}
		\PalIndexEntry{16-23}{Paper}
		\PalIndexEntry{24-31}{Bright paper}
	\end{PalIndexTable}

	Border is taken from paper colours.

	
	\vspace*{1em}
	\textbf{ULA+}

	\begin{PalIndexTable}
		\PalIndexEntry{0-64}{Ink}
	\end{PalIndexTable}

	Paper and border are taken from \PortLink{Transparency Colour Fallback Register}{4A}.


	\vspace*{1em}
	\textbf{ULANext normal mode}

	\begin{PalIndexTable}
		\PalIndexEntry{~~0-127}{Ink (only a subset)}
		\PalIndexEntry{128-255}{Paper (only a subset)}
	\end{PalIndexTable}

	Border is taken from paper colours. The number of active indices depends on the number of attribute bits assigned to ink and paper out of the attribute byte by \PortLink{Enhanced ULA Ink Colour Mask}{42}.


	\vspace*{1em}
	\textbf{ULANext full-ink mode}

	\begin{PalIndexTable}
		\PalIndexEntry{0-255}{Ink}
	\end{PalIndexTable}

	Paper and border are taken from \PortLink{Transparency Colour Fallback Register}{4A}.

\end{quote}


\pagebreak
\subsubsection{\PortTitle{Palette Value Register}{41}}

\begin{NextPort}
	\PortBits{7-0}
		\PortDesc{Reads or writes 8-bit colour data}
\end{NextPort}

Format is:

\begin{BitTableByte}
	\BitSmall{$R_2$} &
		\BitSmall{$R_1$} &
		\BitSmall{$R_0$} &
		\BitSmall{$G_2$} &
		\BitSmall{$G_1$} &
		\BitSmall{$G_0$} &
		\BitSmall{$B_2$} &
		\BitSmall{$B_1$} \\
	\hline
	\BitStartMulti{3}{Red} &
		\BitMulti{3}{Green} &
		\BitMulti{2}{Blue} \\
\end{BitTableByte}

Least significant bit of blue is set to {\tt OR} between $B_2$ and $B_1$.

Writing the value will automatically increment index in \PortLink{Palette Index Register}{40}, if auto-increment is enabled in \PortLink{Enhanced ULA Control Register}{43}. Read doesn't auto-increment index.


\subsubsection{\PortTitle{Enhanced ULA Ink Colour Mask}{42}}

\begin{NextPort}
	\PortBits{7-0}
		\PortDesc{The number for last ink colour entry in the palette. Only used when ULANext mode is enabled (see \PortLink{Enhanced ULA Control Register}{43}). Only the following values are allowed, harware behavior is unpredictable for other values:}
		\PortDescOnly{
			\begin{PortBitConfig}
				\PortBitLine{1}{Ink and paper only use 1 colour each on indices {\tt 0} and {\tt 128} respectively}
				\PortBitLine{3}{Ink and paper use 4 colours each, on indices {\tt 0}-{\tt 3} and {\tt 128}-{\tt 131}}
				\PortBitLine{7}{Ink and paper use 8 colours each, on indices {\tt 0}-{\tt 7} and {\tt 128}-{\tt 135}}
				\PortBitLine{15}{Ink and paper use 16 colours each, on indices {\tt 0}-{\tt 15} and {\tt 128}-{\tt 143}}
				\PortBitLine{31}{Ink and paper use 32 colours each, on indices {\tt 0}-{\tt 31} and {\tt 128}-{\tt 159}}
				\PortBitLine{63}{Ink and paper use 64 colours each, on indices {\tt 0}-{\tt 63} and {\tt 128}-{\tt 191}}
				\PortBitLine{127}{Ink and paper use 128 colours each, on indices {\tt 0}-{\tt 127} and {\tt 128}-{\tt 255}}
				\PortBitLine{255}{Enables full-ink colour mode where all indices are ink. In this mode paper and border are taken from \PortLink{Transparency Colour Fallback Register}{4A}}
			\end{PortBitConfig}
		}
		\PortDescOnly{Default value is {\tt 7} for core 3.0 and later, {\tt 15} for older cores.}
\end{NextPort}


\subsubsection{\PortTitle{Enhanced ULA Control Register}{43}}

\begin{NextPort}
	\PortBits{7}
		\PortDesc{{\tt 1} to disable palette index auto-increment, {\tt 0} to enable}
	\PortBits{6-4}
		\PortDesc{Selects palette for read or write}
		\PortDescOnly{
			\begin{PortBitConfig}
				\PortBitLine{000}{ULA first palette}
				\PortBitLine{100}{ULA second palette}
				\PortBitLine{001}{Layer 2 first palette}
				\PortBitLine{101}{Layer 2 second palette}
				\PortBitLine{010}{Sprites first palette}
				\PortBitLine{110}{Sprites second palette}
				\PortBitLine{011}{Tilemap first palette}
				\PortBitLine{111}{Tilemap second palette}
			\end{PortBitConfig}
		}
	\PortBits{3}
		\PortDesc{Selects active Sprites palette ({\tt 0} = first palette, {\tt 1} = second palette)}
	\PortBits{2}
		\PortDesc{Selects active Layer 2 palette ({\tt 0} = first palette, {\tt 1} = second palette)}
	\PortBits{1}
		\PortDesc{Selects active ULA palette ({\tt 0} = first palette, {\tt 1} = second palette)}
	\PortBits{0}
		\PortDesc{Enables ULANext mode if {\tt 1} ({\tt 0} after reset)}
\end{NextPort}

Write will also reset the index of \PortLink{Enhanced ULA Palette Extension}{44} so next write there will be considered as first byte of first colour.


\subsubsection{\PortTitle{Enhanced ULA Palette Extension}{44}}

\begin{NextPort}
	\PortBits{7-0}
		\PortDesc{Reads or writes 9-bit colour definition}
\end{NextPort}

Two consequtive writes are needed:

\begin{multicols}{2}
	First write:
	
	\begin{BitTableByte}[C{1em}]
		\BitSmall{$R_2$} &
			\BitSmall{$R_1$} &
			\BitSmall{$R_0$} &
			\BitSmall{$G_2$} &
			\BitSmall{$G_1$} &
			\BitSmall{$G_0$} &
			\BitSmall{$B_2$} &
			\BitSmall{$B_1$} \\
		\hline
		\BitStartMulti{3}{Red} &
			\BitMulti{3}{Green} &
			\BitMulti{2}{Blue} \\
	\end{BitTableByte}

	\columnbreak

	Second write:

	\begin{BitTableByte}[C{1em}]
		\BitSmall{$P_r$} &
			\BitMulti{6}{-} &
			\BitSmall{$B_0$} \\
		\hline
		\BitSmall{$L2$} &
			\BitMulti{6}{Reserved, set to {\tt 0}} &
			\BitSmall{B} \\
	\end{BitTableByte}

\end{multicols}

Bit 7 of the second write must be {\tt 0} except for Layer 2 palettes where it specifies colour priority. If set to {\tt 1}, then the colour will always be on top, above all other layers, regardless of priority set with \PortLink{Sprite and Layers System Register}{15}. So if you need exactly the same colour with priority and non-priority, you will need to set the same data twice, to different indexes, once with priority bit {\tt 1} and then with {\tt 0}.

After the second write palette colour index in \PortLink{Palette Index Register}{40} is automatically increment, if auto-increment is enabled in \PortLink{Enhanced ULA Control Register}{43}.

Note: reading will always return the second byte of the colour (least significant bit of blue) and will not auto-increment index. You can read {\tt RRRGGGBB} part with \PortLink{Palette Value Register}{41}.


\subsubsection{\PortTitle{Transparency Colour Fallback Register}{4A}}

\begin{NextPort}
	\PortBits{7-0}
		\PortDesc{8-bit colour to be used when all layers contain transparent pixel. Format is {\tt RRRGGGBB}}
\end{NextPort}

This colour is also used for paper and border when ULANext full-ink mode is enabled - see \PortLink{Enhanced ULA Ink Colour Mask}{42}.


\pagebreak
